Make any stupid plot you want

### 2021-01-19 draft/sections/math.tex
more on fibers + monoids

\subsubsection{Variable Field in F }
\label{sec:data_fiber}
The fiber is a topological space containing the set of all observations that could possible be in the dataset, even values that do not actually exist in the dataset. For example, lets construct a fiber bundle that is just the temperature column from GHCN. Then the fiber $F$ is the range of physically possible temperature values $\{tp_0, \ldots, tp_n\} \subset \mathbb{R}$. Lets rename this
\begin{equation}
F_0 = \{tp_0, \ldots, tp_n\} 
\end{equation}
so that more variables can be added to the fiber. Then lets add the time from GHCN $F_1 =\{t_0, \ldots, t_n\}$
\begin{equation}
F =  \{tp_0, \ldots, tp_n\}  \times \{ti_0, \ldots, ti_n\} 
\end{equation}

The elements in a given fiber set do not need to be scalers; for example the location can be encoded as the set of tuples $(\textit{lat},\textit{lon})$.  We add the location to the fiber 
\begin{equation}
    F = \{tp_0, \ldots, tp_n\}  \times \{ti_0, \ldots, ti_n\} \times 
    \{(la_0, ln_0),\ldots, (la_j, ln_k), \ldots (la_n,ln_n)\}
\end{equation}
such that one record $r$ in the fiber would have the form $(tp, t, (la,ln))$. We also add the variable name and measurement scale $M$ monoid to the fiber $F_i$ such that 
\begin{align}
    F_0 &= (\textit{temperature}, \{tp_0, \ldots, tp_n\}, M_0)\\
    F_1 &= (\textit{time}, \{ti_0, \ldots, ti_n\}, M_1)\\
    F_2 &= (\textit{location},   \{(la_0, ln_0),\ldots, (la_j, ln_k), \ldots (la_n,ln_n)\}, M_2)
\end{align}
 or more generally (name, set, monoid)
 \begin{equation}
    F_i = (\textit{name}, \{x_{i}, \ldots, x_{n}\}, M_{i})
 \end{equation}
as described in Spivak's categorical formulation of databases \cite{spivakSIMPLICIALDATABASES}. We describe $M$ in more detail in section~\ref{sec:data_monoid}. We can keep growing the fiber
\begin{equation}
F= F_{0} \times \ldots \times F_{n}
\end{equation}
while also retaining the individual sets $F_{i}$ that we will transform into visual characteristics of the graphic.

\subsubsection{Measurement Type}
\label{sec:data_monoid}
%% maybe move somewhere else w/ some more motivation
%% do this section in terms of $M0, F_0$
An important property of data fields are the measurement type, which we generalize to identifying the left monoid action such that the monoid actions $M$ 
\begin{equation}
M = M_{0} \times \ldots\times M_{i}
\end{equation}
is the cartesion cross product of monoids applied to each $F_{i}\in F$. A monoid \cite{Monoid2021} $M$ is a set with an associative binary operator $\ast:M \times M\rightarrow M$. A monoid has an identity element $e\in M$ such that $e\ast a= a \ast e = a$ for all $a \in M$. A left monoid action \cite{SemigroupAction2021,ActionNLab} of $M$ is a set $F$ with an action $\bullet: M\times F \rightarrow F$ with the properties:
\begin{description}
    \item[associativity] $\text{for all } m,t \in M \text{ and } x\in F, m\bullet(t\bullet x) = (m\bullet t) \bullet x$
    \item[identity] $\text{for all } x\in F, e\in M,  e\bullet x = x$ 
\end{description}

@@

%%close out the fiber bundle section with the cartesian products, which gives us heterogenous data , univariate to multivariate. We can decompose the fiber bundle to components bundles \pi:E_1\rightarrow K, .... \pi: E_n\rightarrow K
\subsubsection{Connectivity: Base Space} 
\label{sec:data_base}
%% flows out of the fiber section into, now gluing observations together into a thing
%% can then put functions on edges/planes/etc so we have true continouity
We model data as a fiber bundle ($E$, $K$, $\pi$, $F$)
\begin{equation}
    \label{eq:fiber_bundle}
    \begin{tikzcd}
        F \arrow[r, hook] & E \arrow[r, "\pi"] & K
    \end{tikzcd}
\end{equation}

@@

\subsubsection{Data: Section}
\label{sec:data_section}
%%% visual variable encoding in a consistent/atomic way in a complex structure 
\label{sec:section_data}
The sections of the fiber bundles are the instances of the data. They are the functions that map from a point $k \in K$ to a set of values in $F$. For example, if we have a database, the section is the table that yields a row given an index. The section $\tau$ is the mapping from base space to total space $\tau: K\rightarrow E$ 
\begin{equation}
    \label{eq:section}
    \begin{tikzcd}
        F \arrow[r, hook] & E \arrow[d, "\pi"']           \\
        & K \arrow[u, "\tau"', bend right]
    \end{tikzcd}
\end{equation}

@@

\subsubsection{Case Study: Temperature}
\textcolor{red}{Moved \& walked through b/c was getting clunky to not have terms yet}\\

>>>>

\subsubsection{Variables: Fiber Space $F$}
\label{sec:data_fiber}
The fiber is a topological space containing the set of all observations that could possible be in the dataset, even values that do not actually exist in the dataset. For example, a fiber $F$ that is the temperature variable from GHCN is the range of physically possible temperature values $\{tp_0, \ldots, tp_n\} \subset \mathbb{R}$. Lets rename this
\begin{equation}
F_0 = \{tp_0, \ldots, tp_n\} 
\end{equation}
so that more variables can be added to the fiber. Then lets add the time from GHCN $F_1 =\{t_0, \ldots, t_n\}$
\begin{equation}
F =  \{tp_0, \ldots, tp_n\}  \times \{ti_0, \ldots, ti_n\} 
\end{equation}

The elements in a given fiber set do not need to be scalers; for example the location can be encoded as the set of tuples $(\textit{lat},\textit{lon})$.  We add the location to the fiber 
\begin{equation}
    F = \{tp_0, \ldots, tp_n\}  \times \{ti_0, \ldots, ti_n\} \times 
    \{(la_0, ln_0),\ldots, (la_j, ln_k), \ldots (la_n,ln_n)\}
\end{equation}
such that one record $r$ in the fiber would have the form $(tp, t, (la,ln))$. We also add the variable name and measurement scale $M$ monoid actions to the fiber $F_i$ such that 
\begin{align}
    F_0 &= (\textit{temperature}, \{tp_0, \ldots, tp_n\}, M_0)\\
    F_1 &= (\textit{time}, \{ti_0, \ldots, ti_n\}, M_1)\\
    F_2 &= (\textit{location},   \{(la_0, ln_0),\ldots, (la_j, ln_k), \ldots (la_n,ln_n)\}, M_2)
\end{align}
 or more generally (name, set, monoid)
 \begin{equation}
    F_i = (\textit{name}, \{x_{i}, \ldots, x_{n}\}, M_{i})
 \end{equation}
as described in Spivak's categorical formulation of databases \cite{spivakSIMPLICIALDATABASES}. We describe $M$ in more detail in section~\ref{sec:data_monoid}. We can keep growing the fiber
\begin{equation}
F= F_{0} \times \ldots \times F_{n}
\end{equation}
and identify the elements in the record by the variable name. We will often be working with the decoupled fiber $F_i$ rather than the total fiber $F$.

\subsubsection{Measurement Scale: Monoid Actions}
\label{sec:data_monoid}
The monoid actions $M_i$ is the set of transformations on $F_i$ that we expect to be preserved in the visual transform. For example, the temperature $F_0$ with units $\degree C$ 
Steven's measurement scales define group actions \cite{stevensTheoryScalesMeasurement1946, leaFormalizationMeasurementScale}

%% maybe move somewhere else w/ some more motivation
%% do this section in terms of $M0, F_0$
An important property of data fields are the measurement type, which we generalize to identifying the left monoid action such that the monoid actions $M$ 
\begin{equation}
M = M_{0} \times \ldots\times M_{i}
\end{equation}
is the cartesion cross product of monoids applied to each $F_{i}\in F$. A monoid \cite{Monoid2021} $M$ is a set with an associative binary operator $\ast:M \times M\rightarrow M$. A monoid has an identity element $e\in M$ such that $e\ast a= a \ast e = a$ for all $a \in M$. A left monoid action \cite{SemigroupAction2021,ActionNLab} of $M$ is a set $F$ with an action $\bullet: M\times F \rightarrow F$ with the properties:
\begin{description}
    \item[associativity] $\text{for all } m,t \in M \text{ and } x\in F, m\bullet(t\bullet x) = (m\bullet t) \bullet x$
    \item[identity] $\text{for all } x\in F, e\in M,  e\bullet x = x$ 
\end{description}

@@

%%close out the fiber bundle section with the cartesian products, which gives us heterogenous data , univariate to multivariate. We can decompose the fiber bundle to components bundles \pi:E_1\rightarrow K, .... \pi: E_n\rightarrow K
\subsubsection{Connectivity: Base Space $K$} 
\label{sec:data_base}
%% flows out of the fiber section into, now gluing observations together into a thing
%% can then put functions on edges/planes/etc so we have true continouity
We model data as a fiber bundle ($E$, $K$, $\pi$, $F$)
\begin{equation}
    \label{eq:fiber_bundle}
    \begin{tikzcd}
        F \arrow[r, hook] & E \arrow[r, "\pi"] & K
    \end{tikzcd}
\end{equation}

@@

\subsubsection{Values: Sections $\tau$}
\label{sec:data_section}
%%% visual variable encoding in a consistent/atomic way in a complex structure 
\label{sec:section_data}
The sections of the fiber bundles are the instances of the data. They are the functions that map from a point $k \in K$ to a set of values in $F$. For example, if we have a database, the section is the table that yields a row given an index. The section $\tau$ is the mapping from base space to total space $\tau: K\rightarrow E$ 
\begin{equation}
    \label{eq:section}
    \begin{tikzcd}
        F \arrow[r, hook] & E \arrow[d, "\pi"']           \\
        & K \arrow[u, "\tau"', bend right]
    \end{tikzcd}
\end{equation}

@@

\subsubsection{Case Study: GHCN}
\textcolor{red}{Moved \& walked through b/c was getting clunky to not have terms yet}\\
